---
title: "Launch NicheNet for secreted Genes in Fibroblasts-FAPs"
author: "Maria Kondili"
date: "18/07/2022"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(AnnotationDbi))
suppressPackageStartupMessages(library("org.Hs.eg.db"))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(nichenetr)) ## devtools::install_github("saeyslab/nichenetr")
library(Seurat)

```


```{r}

work_dir <- "/projects/single_cell_skeletal_muscle/DeMicheli/human/"

secretome_dir <- paste0(work_dir, "Analyse_NormData_with_Seurat/Secretome/")

out_dir <- paste0(secretome_dir,"NicheNet_analysis/")
```

###### Read Input Genes = Secreted

```{r}
secreted_fibro_genes <- read_delim(paste0(secretome_dir,"markers_fibro/109_Common_Fibroblasts_Secreted_Genes.tsv"),
                                        delim="\n",col_names=T)

secreted_fibro_genes <- secreted_fibro_genes$GeneName
```

##### Define Background-Geneset -> Read All genes of S.C.dataset-human

> metadata:

```{r}
metadata <- read.table(paste0(work_dir,"GSE143704/DeMicheli_HumanMuscleAtlas_metadata.txt"),
                       sep="\t",header=TRUE,row.names=1)

metadata <- metadata[,c(1:6)]

```

>normalised count data  

```{r read_sc_data }
data <- read.table(paste0(work_dir,"GSE143704/DeMicheli_HumanMuscleAtlas_normalizeddata.txt"),
                        sep="\t",header=TRUE,row.names=1)

#data    <- readRDS(paste0(work_dir,"Analyse_NormData_with_Seurat/norm_expr_data.rds"))

srat_data <- CreateSeuratObject(counts = data,
                                 project = "human_muscle",
                                 meta.data = metadata,
                                 min.cells = 3,
                                 min.features = 200)

srat_data <- srat_data %>% ScaleData(features = rownames(srat_data)) %>%
             FindVariableFeatures(selection.method = "vst",nfeatures = 2000) %>%
             RunPCA(features = VariableFeatures(object = .), verbose=TRUE)

srat_data <- FindNeighbors(srat_data, dims = 1:30)

srat_data <- FindClusters(srat_data, resolution = 0.4)

srat_data <- SetIdent(srat_data,value = "cell_annotation")

# saveRDS(srat_data,paste0(work_dir,"Analyse_NormData_with_Seurat/srat_data.rds"))

```

Create the background to be used : All expressed genes  

```{r}

background <- rownames(srat_data@assays$RNA@counts)

```

### NicheNet’s ligand-target prior model

```{r}
httr::timeout(100)
ligand_target_matrix <- readRDS(url("https://zenodo.org/record/3260758/files/ligand_target_matrix.rds"))

ligand_target_matrix %>% glimpse

```

###  NicheNet's Ligand-Receptor Data Source

```{r}
httr::timeout(100)
lr_network <- readRDS(url("https://zenodo.org/record/3260758/files/lr_network.rds"))

head(lr_network)
```

Unique Genes as RECEPTORS found in the network:

```{r}
lr_network$to %>% as.factor %>% levels() %>% length  
```


```{r}
ligands <- lr_network %>%  pull(from) %>%unique()

ligands %>% length
```

### Find Ligands in our Input Dataset of Secreted Genes of FAPs

```{r}

expressed_ligands <- intersect(ligands,secreted_fibro_genes)
expressed_ligands

```

##### Find Receptors in receiver-cells :

1 / Extract Single-Cell expr. of FAPs, Endoth, MuSc, Mature Skeletal.Muscle, Macrophages

```{r}

## extract 3 sub-clusters of Fibroblasts-FAPs
samples_Fibro1 <- rownames(subset(metadata,cell_annotation=="COL1A1+ Fibroblasts") )
samples_Fibro2 <- rownames(subset(metadata,cell_annotation=="DCN+ GSN+ MYOC+ Fibroblasts") )
samples_Fibro3 <- rownames(subset(metadata,cell_annotation=="FBN1+ MFAP5+ CD55+ Fibroblasts") )

samples_Fibro <- unique(c(samples_Fibro1, samples_Fibro2, samples_Fibro3))

## extract samples of other cell-types : Endothelial, Stem-cells,Mature sk.mu, Macrophages
samples_endoth <- rownames(subset(metadata,cell_annotation %in%
                                  c("ICAM1+ SELE+ VCAM1+ Endothelial",
                                    "CLDN5+ PECAM1+ Endothelial")))

samples_mature_skmu <- rownames(subset(metadata,cell_annotation == "ACTA1+ Mature skeletal muscle"))

samples_mature_skmu %>% length()
samples_MuSC <-  rownames(subset(metadata,cell_annotation %in%
                                   c("PAX7+ DLK1+ MuSCs and progenitors",
                                    "PAX7low MYF5+ MuSCs and progenitors")))
samples_MuSC %>% length
samples_macroph <- rownames(subset(metadata,cell_annotation %in%
                                     c("S100A9+ LYZ+ Inflammatory macrophages",
                                       "C1QA+ CD74+ Macrophages")))

samples_macroph %>% length
```

##### Extract GeneNames expressed at each cell-type :

 Keep only genes with expr > 0

```{r}
##> Counts only of these samples -> Find Genes that are expressed in each cell-type

subset_celltype_and_genes <- function(srat_data,ct_samples){

  #1. choose samples
  srat_counts <- srat_data@assays$RNA@counts[, ct_samples]

  #2. filter lines for genes with counts >0
  srat_counts_filt <- srat_counts[which(rowSums(srat_counts) >0 ),]

  genes_ct <- rownames(srat_counts_filt)
  return(genes_ct)

}


#srat_counts_fibro  <- srat_data@assays$RNA@counts[,c(samples_Fibro1,samples_Fibro2,samples_Fibro3)]
fibro_genes   <- subset_celltype_and_genes(srat_data,
                                          c(samples_Fibro1,samples_Fibro2,samples_Fibro3))

endoth_genes    <- subset_celltype_and_genes(srat_data,samples_endoth)

mature_mu_genes <- subset_celltype_and_genes(srat_data,samples_mature_skmu)

MuSC_genes      <-  subset_celltype_and_genes(srat_data,samples_MuSC)

macroph_genes   <- subset_celltype_and_genes(srat_data,samples_macroph)


```

2/ Find which genes from the above cell-types can be regulated :

Potential Receptors :

- fibroblasts (3 sub-clusters)

- endothelial cells

- mature_skeletal muscle cells

- Muscle stem-cells and progenitors

- macrophages

```{r}

receptors <- lr_network %>% pull(to) %>% unique()

expr_receptors_fibro <- intersect(receptors, fibro_genes)
expr_receptors_endoth <- intersect(receptors, endoth_genes)
expr_receptors_mature_mu <- intersect(receptors, mature_mu_genes)
expr_receptors_MuSC <- intersect(receptors, MuSC_genes)
expr_receptors_macroph <- intersect(receptors, macroph_genes)

glue::glue("There are {expr_receptors_fibro %>% length} Receptors ,among genes expressed in Fibroblast Cells.")
glue::glue("There are {expr_receptors_endoth %>% length} Receptors ,among genes expressed in Endothelial Cells.")
glue::glue("There are {expr_receptors_mature_mu %>% length} Receptors ,among genes expressed in Mature Skel.muscle cells")
glue::glue("There are {expr_receptors_MuSC %>% length} Receptors ,among genes expressed in Stem Cells & Progenitors.")
glue::glue("There are {expr_receptors_macroph %>% length} Receptors ,among genes expressed in Macrophages.")


```


```{r}

find_ligand_targets_in_SC_FAPs <- function(lr_network,
                                           background,
                                           secreted_genes, # e.g fibro_genes
                                           expressed_ligands,
                                           expressed_receptors,
                                           ligand_target_matrix,
                                           cell_type){

  ##### Filter Ligands and receptors of Nichenet with only expressed-ones in my dataset
  lr_network_expressed <- lr_network %>%
                          filter(from %in% expressed_ligands & to %in% expressed_receptors)

  potential_ligands <- lr_network_expressed %>%
                        pull(from) %>%
                        unique()

  glue::glue("Potential ligands are :\n {potential_ligands}")

  ##### Calculate the ligand activity of each ligand


  ligand_activities <- predict_ligand_activities(geneset = secreted_genes,
                                                background_expressed_genes = background,
                                                ligand_target_matrix = ligand_target_matrix,
                                                potential_ligands = potential_ligands)

  ##> To prioritize by Pearson-Correlation-Coefficient :
  ##> ligand_activities  %>% arrange(-pearson)


  best_upstream_ligands <- ligand_activities %>%
                            top_n(20, pearson) %>%
                            arrange(-pearson) %>%
                            pull(test_ligand)



  active_ligand_target_links_df <- best_upstream_ligands %>%
                                    lapply(get_weighted_ligand_target_links,
                                           geneset = secreted_genes ,
                                           ligand_target_matrix = ligand_target_matrix,n = 250) %>%
                                    bind_rows()

  # active_ligand_target_links_df %>% head
  active_ligand_target_links_df %>% nrow


  active_ligand_target_links <- prepare_ligand_target_visualization(
                                ligand_target_df     = active_ligand_target_links_df,
                                ligand_target_matrix = ligand_target_matrix,
                                cutoff = 0.25)
  # cut-off :if >>0.25 ,only strong L-Targets kept, if <<0.25, more links appear

  active_ligand_target_links %>% head


  order_ligands <- intersect(best_upstream_ligands, colnames(active_ligand_target_links)) %>% rev()
  order_targets <- active_ligand_target_links_df$target %>% unique()


 #### Create square Matrix : ligands x receptors

  vis_ligand_target <- active_ligand_target_links[order_targets,order_ligands] %>% t()


  ###### Heatmap of Ligands with Potentially regulated genes

  p_ligand_target_network <-  vis_ligand_target %>%
                              make_heatmap_ggplot("Secreted-FAPs ligands",
                                                  paste0("Potentially regulated genes in ",cell_type),
                                                  color = "purple",
                                                  legend_position = "top",
                                                  x_axis_position = "top",
                                                  legend_title = "Regulatory potential") +
                              scale_fill_gradient2(low = "whitesmoke",  
                                                   high = "purple",
                                                   breaks = c(0,0.005,0.01)) +
                              theme(axis.text.x = element_text(face = "italic", size=9)) +
                              ggtitle(cell_type)


  return(p_ligand_target_network)
}

```


```{r}

#pdf("FAPs_Ligands_and_Potential_Receptors_in_other_cell_types.pdf",onefile=T)
find_ligand_targets_in_SC_FAPs(lr_network,
                              background,
                              secreted_fibro_genes,
                              expressed_ligands,
                              expr_receptors_fibro,
                              ligand_target_matrix,
                              "Fibroblasts" )  %>% print(.)

##! For interactive exploration of plot use plotly:
#ggplotly(plot_lr_fibro)

# ggsave(plot_lr_fibro, filename="FAPs_Ligands_and_Potential_Receptors_Fibroblasts.pdf",
#        device = cairo_pdf, width = 5, height = 4)


find_ligand_targets_in_SC_FAPs(lr_network,
                                  background,
                                  secreted_fibro_genes,
                                  expressed_ligands,
                                  expr_receptors_endoth,
                                  ligand_target_matrix,
                                  "Endothelial_Cells")  %>% print(.)

find_ligand_targets_in_SC_FAPs(lr_network,
                                  background,
                                  secreted_fibro_genes,
                                  expressed_ligands,
                                  expr_receptors_mature_mu,
                                  ligand_target_matrix,
                                  "Mature_sk_Mu_Cells" ) %>%  print(.)



find_ligand_targets_in_SC_FAPs(lr_network,
                                  background,
                                  secreted_fibro_genes,
                                  expressed_ligands,
                                  expr_receptors_MuSC,
                                  ligand_target_matrix,
                                  "MuSCs_and_progenitors") %>% print(.)


find_ligand_targets_in_SC_FAPs(lr_network,
                                  background,
                                  secreted_fibro_genes,
                                  expressed_ligands,
                                  expr_receptors_macroph,
                                  ligand_target_matrix,
                                  "Macrophages") %>% print(.)


```

#### LIGAND-RECEPTORS PAIRS

```{r}

get_best_ligands <- function(lr_network,background,
                             ligand_target_matrix,
                             secreted_genes,expressed_receptors){

  lr_network_expressed <- lr_network %>%
                          filter(from %in% secreted_genes & to %in% expressed_receptors)

  potential_ligands <- lr_network_expressed %>% pull(from) %>% unique()

  glue::glue("Potential ligands are  : {potential_ligands}")

  ##### Calculate the ligand activity of each ligand
  ligand_activities <- predict_ligand_activities(geneset = secreted_genes,
                                                background_expressed_genes = background,
                                                ligand_target_matrix = ligand_target_matrix,
                                                potential_ligands = potential_ligands)

  ##> To prioritize by Pearson-Correlation-Coefficient :
  ##> ligand_activities  %>% arrange(-pearson)

  best_upstream_ligands <- ligand_activities %>%
                            top_n(30, pearson) %>%
                            arrange(-pearson) %>%
                            pull(test_ligand)

  return(best_upstream_ligands)
}

```

#### Define Best Ligands to be considered for L-R network (for each cell-type)

```{r}

best_upstream_ligands_fibro <- get_best_ligands(lr_network,
                                          background,
                                          ligand_target_matrix,
                                          secreted_fibro_genes,
                                          expr_receptors_fibro)
best_upstream_ligands_fibro

best_upstream_ligands_endoth <- get_best_ligands(lr_network,
                                          background,
                                          ligand_target_matrix,
                                          secreted_fibro_genes,
                                          expr_receptors_endoth)
best_upstream_ligands_endoth

best_upstream_ligands_macrof <- get_best_ligands(lr_network,
                                          background,
                                          ligand_target_matrix,
                                          secreted_fibro_genes,
                                          expr_receptors_macroph)
best_upstream_ligands_macrof


best_upstream_ligands_skmu <- get_best_ligands(lr_network,
                                          background,
                                          ligand_target_matrix,
                                          secreted_fibro_genes,
                                          expr_receptors_mature_mu)
best_upstream_ligands_skmu

best_upstream_ligands_muSc <- get_best_ligands(lr_network,
                                          background,
                                          ligand_target_matrix,
                                          secreted_fibro_genes,
                                          expr_receptors_MuSC)

best_upstream_ligands_muSc
```

###### Find Best Ligand-Receptor Pairs

```{r}

get_best_ligand_receptor_pairs <- function(lr_network,best_upstream_ligands, expr_receptors, cell_type){

  ## Now we get the ligand-receptor network of the top-ranked ligands
  ##> To replace expr_receptors --> expr_receptors_<other_celltypes>

  lr_network_top <- lr_network %>%
                    filter(from %in% best_upstream_ligands & to %in% expr_receptors) %>%
                    distinct(from,to)

  #>"distinct" fun: remove "source" and "database" columns, keep unique pairs(removes duplic.rows)

  best_upstream_receptors <- lr_network_top %>% pull(to) %>% unique()

  ## Get the weights of the ligand-receptor interactions as used in the NicheNet model

  httr::timeout(100)
  weighted_networks = readRDS(url("https://zenodo.org/record/3260758/files/weighted_networks.rds"))

  ## Keep only best L-R from the Network :
  lr_network_top_df <- weighted_networks$lr_sig %>%
                       filter(from %in% best_upstream_ligands & to %in% best_upstream_receptors )

  ##> create square table with "from"-genes in columns ,via "spread()" function :
  lr_network_top_df <- lr_network_top_df %>% spread("from","weight",fill = 0)

  lr_network_top_matrix <- lr_network_top_df %>%
                           dplyr::select(-to) %>% as.matrix() %>%
                           magrittr::set_rownames(lr_network_top_df$to)


  ## Hierarchical Clustering to order the ligands and receptors:
  dist_receptors <- dist(lr_network_top_matrix, method = "binary")
  hclust_receptors <- hclust(dist_receptors, method = "complete" ) #method="median" , "median" , "ward.D2"
  order_receptors_hcl <- hclust_receptors$labels[hclust_receptors$order]

  dist_ligands <- dist(lr_network_top_matrix %>% t(), method = "binary")
  hclust_ligands <- hclust(dist_ligands, method = "complete")
  order_ligands_hcl <- hclust_ligands$labels[hclust_ligands$order]


  vis_ligand_receptor_network <- lr_network_top_matrix[order_receptors_hcl, order_ligands_hcl]

  p_ligand_receptor_network <- vis_ligand_receptor_network %>% t() %>%
                               make_heatmap_ggplot("FAPs secreted ligands",
                                                  paste0("Receptors expressed by ", cell_type),
                                                  color = "mediumvioletred",
                                                  x_axis_position = "top",
                                                  legend_title = "Prior interaction potential")

  return(p_ligand_receptor_network)
}

```


```{r}
#pdf("FAPs_Ligands_and_Receptors_of_other_cell_types_heatmaps.pdf")

get_best_ligand_receptor_pairs(lr_network,
                               best_upstream_ligands_fibro,
                               expr_receptors_fibro,
                               "Fibroblasts_3subclusters") %>% print(.)

get_best_ligand_receptor_pairs(lr_network,
                               best_upstream_ligands_skmu,
                               expr_receptors_mature_mu,
                               "Mature SkMuscle") %>% print(.)


get_best_ligand_receptor_pairs(lr_network,
                               best_upstream_ligands_endoth,
                               expr_receptors_endoth,
                               "Endothelial Cells") %>% print(.)
```

Many Receptors are common in every graph from different cell-type ...!
Let's check globally what happens with these receptors found :


#### Show Common Receptors among cell-types

```{r}

### each one with fibroblasts

intersect(expr_receptors_fibro,expr_receptors_endoth) %>% length
intersect(expr_receptors_mature_mu,expr_receptors_fibro) %>% length
intersect(expr_receptors_MuSC,expr_receptors_fibro) %>% length
intersect(expr_receptors_macroph,expr_receptors_fibro) %>% length

### among all cell-types :
Reduce(intersect,
       list(expr_receptors_macroph,
            expr_receptors_fibro,
            expr_receptors_MuSC,
            expr_receptors_mature_mu,
            expr_receptors_endoth)) %>% length

```

##### Visualise Network

based on instructions here :https://github.com/saeyslab/nichenetr/blob/master/vignettes/ligand_target_signaling_path.md
```{r}
sig_network = readRDS(url("https://zenodo.org/record/3260758/files/signaling_network.rds"))
gr_network = readRDS(url("https://zenodo.org/record/3260758/files/gr_network.rds"))

```


```{r}

myligands ="FN1"
mytargets = c("CYP1B1", "GSN","GPNMB") # regulated-genes in fibroblasts

```


```{r}
library(xml2)
library(DiagrammeRsvg) # devtools::install_github('rich-iannone/DiagrammeRsvg')

active_signaling_network = get_ligand_signaling_path(ligand_tf_matrix=ligand_tf_matrix,
                                                     ligands_all=myligands,
                                                     targets_all=mytargets,
                                                     weighted_networks=weighted_networks)

# For better visualization of edge weighτs: normalize edge weights to make them comparable between signaling and gene regulatory interactions
#active_signaling_network_min_max = active_signaling_network

active_signaling_network$sig = active_signaling_network$sig %>%
                                mutate(weight = ((weight-min(weight))/(max(weight)-min(weight))) + 0.75)

active_signaling_network$gr = active_signaling_network$gr %>%
                                mutate(weight = ((weight-min(weight))/(max(weight)-min(weight))) + 0.75)

graph_min_max = diagrammer_format_signaling_graph(signaling_graph_list = active_signaling_network,
                                                  ligands_all = myligands,
                                                  targets_all = mytargets,
                                                  sig_color = "indianred",
                                                  gr_color = "steelblue")

DiagrammeR::render_graph(graph_min_max, layout = "tree")

## To save in file :
##  %>% export_svg() %>% read_xml() %>% write_xml("FN1_to_CYP1B1+GSN_network.svg")

```
