---
title: "Analyse S.C. Oprescu-mouse"
author: "Maria Kondili"
date:   "03/01/2022"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r libraries }
library(Seurat)
library(cerebroApp)
library(dplyr)
library(tidyverse)
library(SingleR)
library(celldex)
library(RColorBrewer)
library(GEOquery) # BiocManager::install("GEOquery")
```

```{r input_data}

work_dir <- "Secretome_in_Single_Cell_RNAseq_skeletal_muscle/Oprescu_mouse/"
setwd(work_dir)

data_dir <- paste0(work_dir,"GSE138826/")
dir.create(data_dir, showWarnings=F)

###
###> If downloaded with GEOquery :
###

dir.create("GSE138826/")
#> get file
regen_file <- getGEOSuppFiles(GEO="GSE138826",filter_regex="_regen_data.rds.gz", baseDir="GSE138826/", fetch_files= TRUE)
#> unzip
regen_file_unz <-  str_remove(rownames(regen_file), ".gz")
unzipped_file <- gunzip(rownames(regen_file),destname=regen_file_unz, remove = FALSE)
## read it
regen_data <- readRDS(unzipped_file)

###
###> If downloaded manually:
###
#regen_data <- readRDS(paste0(work_dir,"GSE138826/GSE138826_regen_data.rds"))

#>> regen_data@data  :'data.frame':	18658 obs. of  53193 variables:
#$RNA : contains the RNA counts
#$SCT: contains the SCT normalized values. Columns are the samples, rows are genes

metadata_table <- data.frame("cell_annot"= regen_data$SCT@varMetadata$metacluster,
                             "cell_ID"   = regen_data$SCT@varMetadata$cellID,
                             "timepoint" = regen_data$SCT@varMetadata$timepoint,
                             row.names   ="cell_ID")

regen_data$SCT@data %>% glimpse

# According Article : re-scaled the data using the SCTransform function
```

```{r create_seurat}
### APPLY SEURAT analysis steps documented in :
### https://www.singlecellcourse.org/scrna-seq-dataset-integration.html

srat_data <- CreateSeuratObject(counts=regen_data$SCT@data ,
                                 project="Oprescu",
                                 assay="RNA",
                                 names.field = 2,
                                 names_delim ="_",
                                 meta.data   = metadata_table)

##NOTE: meta.data should contain cell_names in rows and additional meta-info in columns
##        cell_names in rows should match the column_names of raw_data

## Calculate fractions of Ribosomal Proteins and Mitochondrial genes
srat_data[["percent.mt"]]  <- PercentageFeatureSet(srat_data, pattern="^MT-")
srat_data[["percent.rbp"]] <- PercentageFeatureSet(srat_data, pattern="^RP[SL]")

##From Article:
# cells with more than
# 15% reads mapping to mitochondrial genes, feature counts greater than
#6,000 and more than 60,000 reads were filtered out;

VlnPlot(srat_data, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

```

```{r seurat_pipe}
srat_data <-  srat_data %>%
              ScaleData(features = rownames(srat_data)) %>%
              FindVariableFeatures(selection.method = "vst",nfeatures = 2000) %>%
              RunPCA(features = VariableFeatures(object = .),verbose=TRUE)


DimPlot(srat_data, reduction = "pca", repel=TRUE,label=F)

srat_data <- FindNeighbors(srat_data, dims = 1:10)

srat_data <- FindClusters(srat_data, resolution = 0.4)

##>Run non-linear dimensional reduction
srat_data <- RunUMAP(srat_data, dims = 1:20)

srat_data %>% glimpse

```

```{r plot umap}
### Plot with Cell-Type ID :

dittoSeq::dittoDimPlot(srat_data, var = metadata_table$cell_annot,
                       reduction.use = "umap",
                       do.label = T, main="Oprescu-Mouse-S.C.Clusters",
                       labels.size = 3.5, legend.show = F)

### Plot with Cluster-Number
dittoSeq::dittoDimPlot(srat_data, var = "ident",
                       reduction.use = "umap",
                       do.label = T, main="Mouse S.C.Clusters",
                       size = 0.3)
```

```{r find_marker_genes}
# faps_Markers <- Seurat::FindMarkers(srat_data, ident.1=c(1,3,8,17),
#                                     ident.2 = NULL, logfc.threshold = 0.25,
#                                     only.pos = T,min.cells.feature = 3)

faps_markers <- readr::read_delim("Table_S2_FAP_gene_signature_list_per_subtype.tsv",
                                  delim="\t", col_names=T)

# A tibble: 6 × 7
# Cxcl14_FAPs Dpp4_FAPs Dlk1_FAPs Osr1_FAPs Wisp1_FAPs Activated_FAPs Fibroblasts
# <chr>       <chr>     <chr>     <chr>     <chr>      <chr>          <chr>
# 1 Cxcl14      Pi16      Dlk1      Gsn       Postn      Cxcl5          Meg3
# 2 Smoc2       Igfbp5    Itm2a     Mgp       Csrp2      Prg4           Col3a1
# 3 Ccl11       Igfbp6    Meg3      Dcn       Sfrp2      Timp1          Dlk1
# 4 Gsn         Fbn1      Igf2      Smoc2     Ptn        Mt2            Itm2a
# 5 Dcn         Ugdh      Cdkn1c    Col3a1    Cthrc1     Cxcl1          Col1a1
# 6 Hsd11b1     Cd55      Col3a1    Egr1      H19        Ptx3           Col1a2

```

```{r extract_geneNames}
faps_geneNames <- as.vector(unique(unlist(c(faps_markers[,"Cxcl14_FAPs"],
                    faps_markers[,"Dpp4_FAPs"],
                    faps_markers[,"Dlk1_FAPs"],
                    faps_markers[,"Osr1_FAPs"],
                    faps_markers[,"Wisp1_FAPs"],
                    faps_markers[,"Activated_FAPs"]))))

write_file(paste(faps_geneNames,collapse="\n"),"FAPs_Markers_6subgroups_Oprescu.txt")

```

##### FOCUS ON COUNTS of FAPs clusters:

```{r samples_per_timepoint}
samples_FAPs <- rownames(subset(metadata_table,cell_annot=="FAPs"))
samples_FAPs <- gsub(" ", "", samples_FAPs)
idx_faps_colnames <- which(colnames(srat_data@assays$RNA@counts) %in% samples_FAPs)

faps_counts <- srat_data@assays$RNA@counts[,idx_faps_colnames]
faps_counts <- as.data.frame(faps_counts )

timepoints <- map(colnames(faps_counts),
              ~unlist(strsplit(.x, "_"))[1] ) %>%
              unlist %>% as.factor %>% levels

# timepoints
# [1] "0.5DPI"  "10DPI"  "21DPI"   "2DPI"   "3.5DPI"  "5DPI"  "Noninjured"

NI.cols   <- grep("Noninjured",colnames(faps_counts))
T0_5.cols <- grep("0.5DPI",colnames(faps_counts))
T2.cols   <- grep("2DPI",colnames(faps_counts) )
T3_5.cols <- grep("3.5DPI",colnames(faps_counts))
T5.cols   <- grep("5DPI",colnames(faps_counts))
T10.cols  <- grep("10DPI",colnames(faps_counts))
T21.cols  <- grep("21DPI",colnames(faps_counts))

faps_samples_n_genes <- faps_counts[faps_geneNames, ]

```

#### Calculate Mean-Expr Per Timepoint (mean over columns of same tp)

```{r mean_per_tp}
NI.mean   <- rowMeans(faps_counts[,NI.cols])
T0_5.mean <- rowMeans(faps_counts[,T0_5.cols])
T2.mean   <- rowMeans(faps_counts[,T2.cols])
T3_5.mean <- rowMeans(faps_counts[,T3_5.cols])
T5.mean   <- rowMeans(faps_counts[,T5.cols])
T10.mean  <- rowMeans(faps_counts[,T10.cols])
T21.mean  <- rowMeans(faps_counts[,T21.cols])

###> Create a table of Mean-Expr &  Keep only FAPs-Markers

timepoint_mean_faps <- data.frame("NI"=NI.mean, "T0_5"=T0_5.mean,
                                  "T2"=T2.mean, "T3_5"=T3_5.mean,
                                  "T5"=T5.mean, "T10" =T10.mean,
                                  "T21"=T21.mean)

timepoint_MarkerGenes_Faps <- timepoint_mean_faps[faps_geneNames, ]

timepoint_MarkerGenes_Faps %>% dim
#[1] 215   7
```

```{r prepare_counts_faps_tab}

##> Clean from NAs & apply Z-score:
naked_expr_faps <- timepoint_MarkerGenes_Faps
rownames(naked_expr_faps) <- NULL


calc_z_score <- function(x){
  # x -> column of the data.frame
  (x - mean(x)) / sd(x)
}

z_faps <- t(apply(naked_expr_faps,1, calc_z_score))
na_idx <- which(rowSums(is.na(z_faps)) > 0)
na_idx %>% length
z_faps <- drop_na(as.data.frame(z_faps))
z_faps %>% dim
design <- data.frame("sample"   =colnames(timepoint_mean_faps),
                     "replicate"=c(1,1,1,1,1,1,1),
                     "treatment"= colnames(timepoint_mean_faps))

print(design)

col_data <- data.frame(row.names=design$sample, conditions=design$treatment )
col_data  
```

####
#### K-Means Heatmap : Grouping of genes in boxes
####

```{r heatmap}

library(pheatmap)
p_k <- pheatmap(z_faps,
                color = colorRampPalette(rev(brewer.pal(n=10,name ="RdYlBu")))(100),
                annotation_col = col_data,
                cluster_rows = TRUE,
                kmeans_k = 6,
                clustering_method = "complete",
                clustering_distance_rows="euclidean",
                cluster_cols=FALSE,
                annotation_legend = TRUE,
                annotation_names_col = TRUE,
                show_rownames=TRUE,
                labels_row = NULL,
                angle_col = "0",
                main = "K-means Clustering of FAPs-Marker-Genes through Time")
print(p_k)

```

#### Save GeneNames found in each cluster

```{r faps_data_frame }

fapsMarkers_kmeans_clusters <- data.frame("geneName"= faps_geneNames,
                                          "cluster" = p_k$kmeans$cluster )

fapsMarkers_kmeans_clusters[1:10,]

write_delim(fapsMarkers_kmeans_clusters,
            file="FAPs_Markers_in_6clusters_kmeans_heatmap.tsv",
            delim="\t",col_names=TRUE)

```

#### Count Cell-Number per Cell-Type --> Histogram

```{r cell_number_count}
cell_types <- metadata_table$cell_annot %>% as.factor %>% levels

nb_cell_types <- map(cell_types,  ~ length(rownames(subset(metadata_table,cell_annot==.x ))) )
names(nb_cell_types) <- cell_types

nb_cell_per_type <- data.frame(cbind("cell_type"=names(nb_cell_types), "nb_cells"=nb_cell_types))

write_delim(nb_cell_per_type, file="Nb_Cells_Per_CellType.tsv", delim="\t",col_names=T)

```

#### Split per Timepoint :

```{r}
##> can use srat_data:
#colnames(srat_data@assays$RNA@counts) --> indicate timepoint in suffix
#non_inj_cells <- as_tibble(srat_data@assays$RNA@counts) %>% dplyr::select_if(grepl("Noninjured",colnames(.) )) %>% length()

###> or Metadata:
timepoints_md <- metadata_table$timepoint %>% as.factor  %>% levels
## [1] "0_5dpi"  "10dpi"   "21dpi"   "2dpi"    "3_5dpi"  "5dpi"    "non-inj"


### OR: per timepoint->count cells:
nb_cells_per_ct_per_tp <- list()
for (tp in timepoints_md){
  nb_cells_per_ct_per_tp[[tp]] <- map(cell_types, ~nrow(subset(metadata_table, (cell_annot==.x & timepoint==tp ))) )
  names(nb_cells_per_ct_per_tp[[tp]]) <- cell_types
}

##>Transform in Data.Frame (melted-> for ggplot)

cell_count_per_ct_and_tp_df <- data.frame("count"= unlist(nb_cells_per_ct_per_tp))

cell_count_per_ct_and_tp_df$timepoint <- unlist(map(1:nrow(cell_count_per_ct_and_tp_df),
                                                    ~strsplit(rownames(cell_count_per_ct_and_tp_df), "\\.")[[.x]][1]))

cell_count_per_ct_and_tp_df$cell_type <- unlist(map(1:nrow(cell_count_per_ct_and_tp_df),
                                                ~strsplit(rownames(cell_count_per_ct_and_tp_df), "\\.")[[.x]][2]))

```

#### PLOT histogram of Nb.Cells Per Timepoint

```{r}
subset(cell_count_per_ct_and_tp_df, (timepoint=="10dpi" & cell_type=="APC"))
#  rowID         count timepoint     cell_type
# 10dpi.APC       393     10dpi           APC

##> change Order of Timepoints to be sequential :
cell_count_per_ct_and_tp_df$timepoint <- factor(cell_count_per_ct_and_tp_df$timepoint,
                                         levels=c("non-inj","0_5dpi","2dpi","3_5dpi", "5dpi","10dpi","21dpi" ))

## Count total number of Cells -> to show on bars:

overall_cell_count <-  cell_count_per_ct_and_tp_df %>%
                        dplyr::group_by(timepoint) %>%
                        dplyr::summarise(tot_count = sum(count))


ggplot(cell_count_per_ct_and_tp_df, aes(x=timepoint, y=count, fill=cell_type)) +
      geom_bar(stat="identity", width=0.9, position="stack") +
      ggrepel::geom_label_repel(aes(x=timepoint, y=count,label=count),
                position=position_stack(vjust=0.5)) +
      ggtitle("Cell Count per TimePoint & Cell-Type")

```

####
####  Heatmap only with Secreted
####

```{r}
secretome_dir <- paste0(work_dir,"Secretome/Outcyte/Results/")

outcyte_res <- read_delim(paste0(secretome_dir, "Outcyte_predictions_FAPsMarkers_with_GeneName.tsv"),
                         delim="\t",col_names=T)

## secreted_outcyte_res <- outcyte_res[which(outcyte_res$Prediction %in%  c("Signal-peptide","UPS")),]
secreted_genes <- outcyte_res$GeneSymbol[which(outcyte_res$Prediction %in%  c("Signal-peptide","UPS"))]


pos_secreted <- which(faps_geneNames %in% secreted_genes)


secreted_Faps_expr <- timepoint_MarkerGenes_Faps[pos_secreted , ]
#[1] 215   7


##> Clean from NAs & apply Z-score:
naked_secreted_expr_faps <- secreted_Faps_expr
rownames(naked_secreted_expr_faps) <- NULL


z_faps <- t(apply(naked_secreted_expr_faps,1, calc_z_score))
na_idx <- which(rowSums(is.na(z_faps)) > 0)
na_idx %>% length
z_faps <- drop_na(as.data.frame(z_faps))
z_faps %>% dim
design <- data.frame("sample"   =colnames(timepoint_mean_faps),
                     "replicate"=c(1,1,1,1,1,1,1),
                     "treatment"= colnames(timepoint_mean_faps) )

print(design)

col_data <- data.frame(row.names=design$sample, conditions=design$treatment )

####
#### K-Means Heatmap : Grouping of genes in boxes
####

library(pheatmap)
p_k_secr <- pheatmap(z_faps,
                color = colorRampPalette(rev(brewer.pal(n=10,name ="RdYlBu")))(100),
                annotation_col = col_data,
                cluster_rows = TRUE,
                kmeans_k = 6,
                clustering_method = "complete",
                clustering_distance_rows="euclidean",
                cluster_cols=FALSE,
                annotation_legend = TRUE,
                annotation_names_col = TRUE,
                show_rownames=TRUE,
                labels_row = NULL,
                angle_col = "0",
                main = "K-means Clustering of Secreted-FAPs-Markers through Time")

png("Kmeans_Clustering_k6_Secreted_FAPs_Genes_with_clusterName.png",
    width=700,height=600)
    print(p_k_secr)
dev.off()



###
### Save Names found in each Cluster-Kmeans :
###
SecretedFaps_kmeans_clusters <- data.frame("geneName"= rownames(secreted_Faps_expr),
                                           "cluster" = p_k_secr$kmeans$cluster )

SecretedFaps_kmeans_clusters %>% nrow
##> 157

write_delim(SecretedFaps_kmeans_clusters,
            file="Seurat_Analysis/Secreted_FAPs_Markers_in_6clusters_kmeans_heatmap.tsv",
            delim="\t",col_names=TRUE)

####
#### GO-terms of Secreted-Genes per Cluster
####

source(paste0(work_dir,"functions/create_barplot_with_GOterms_mm.R"))
##> Saves images in folder : Plots/

create_barplot_GOterms.mm(subset(SecretedFaps_kmeans_clusters, cluster == 1)$geneName,
                          gene_format = "SYMBOL",my_ontol="ALL",main="Cluster1-Genes",in_png=T)

create_barplot_GOterms.mm(subset(SecretedFaps_kmeans_clusters, cluster == 2)$geneName,
                          gene_format = "SYMBOL",my_ontol="ALL",main="Cluster2-Genes",in_png=T)

create_barplot_GOterms.mm(subset(SecretedFaps_kmeans_clusters, cluster == 3)$geneName,
                          gene_format = "SYMBOL",my_ontol="ALL",main="Cluster3-Genes",in_png=T)

create_barplot_GOterms.mm(subset(SecretedFaps_kmeans_clusters, cluster == 4)$geneName,
                          gene_format = "SYMBOL",my_ontol="ALL",main="Cluster4-Genes",in_png=T)

create_barplot_GOterms.mm(subset(SecretedFaps_kmeans_clusters, cluster == 5)$geneName,
                          gene_format = "SYMBOL",my_ontol="ALL",main="Cluster5-Genes",in_png=T)

create_barplot_GOterms.mm(subset(SecretedFaps_kmeans_clusters, cluster == 6)$geneName,
                          gene_format = "SYMBOL",my_ontol="ALL",main="Cluster6-Genes",in_png=T)
