---
title: "Analyse S.C.DeMicheli-human"
author: "Maria Kondili"
date:   "25/11/2021"
output: html_document
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```


```{r libraries }
library(Seurat)
suppressPackageStartupMessages(library(tidyverse))
#BiocManager::install('romanhaa/cerebroApp')
library(cerebroApp)
#launchCerebro()

```

##### Read data

```{r data, echo=TRUE,include=TRUE}
##> rownames = 1 : In the rownames are given the gene-symbols

work_dir <-"/shared/projects/single_cell_skeletal_muscle/DeMicheli/human/"

raw_data <- read.table(paste0(work_dir,"GSE143704/DeMicheli_HumanMuscleAtlas_rawdata.txt"),sep="\t",header=TRUE,row.names=1)

metadata <- read.table(paste0(work_dir,"GSE143704/DeMicheli_HumanMuscleAtlas_metadata.txt"),sep="\t",header=TRUE,row.names=1)

metadata <- metadata[,c(1:6)]

srat_data <- CreateSeuratObject(counts = raw_data,
                                project = "human_muscle",
                                meta.data= metadata,
                                min.cells = 3,
                                min.features = 200)

# i:
# min.cells= sets the threshold for genes to only take the genes that are present in at least a specified number of cells.
# min.features= sets the thresholds for cells that express at least a specified number of genes.


```

Note : When reading ^MT- from raw data -> we get more than 40% of Percent_Mitochondrial genes --> The raw data is unfiltered, but the Info in Metadata is Filtered ,with MT% < 20%.

```{r show_percent_mito_etc, echo=T,include=T}
##> Use metadata for the  %Mitoch.genes
## keep count per cell-type :

counts_per_ct <- metadata %>% group_by(cell_annotation) %>% count()

meta_percent_mito <- metadata$percent_mito
names(meta_percent_mito) <- rownames(metadata)
srat_data[["percent_mito"]]  <- meta_percent_mito
srat_data[["percent_rbp"]]   <- PercentageFeatureSet(srat_data, pattern="^RP[SL]")
srat_data[["percent_plt"]]   <- PercentageFeatureSet(srat_data, pattern = "^PPBP")

VlnPlot(srat_data, features = c("nFeature_RNA",
                                "nCount_RNA",
                                "percent_mito",
                                "percent_rbp",
                                "percent_plt"), ncol = 5) +
theme(plot.title = element_text(size=15))

```

But ,if we use the information of MT-genes from raw data we get :
```{r with_Real_Mitoch_Percentage, echo=T,include=T}

srat_data[["percent_mito"]]  <- PercentageFeatureSet(srat_data, pattern = "^MT-")

VlnPlot(srat_data, features = c("nFeature_RNA",
                                "nCount_RNA",
                                "percent_mito",
                                "percent_rbp",
                                "percent_plt"), ncol = 5) +
theme(plot.title = element_text(size=15))

```

##### FILTER data

```{r filter_mito_genes,echo=T,include=T }

FeatureScatter(srat_data, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

x <- subset(srat_data, percent_mito < 20)
count_per_ct_x <- x@meta.data %>% group_by(cell_annotation) %>% count

# Check : n1/n2 of 2 diff. counts per cell-type

srat_data <- subset(srat_data, subset = nCount_RNA > 1000 & percent_mito < 20)
##Could also use:  "& nFeature_RNA < 2500 "

##--> verify mitoch_gene%
VlnPlot(srat_data, features = c("nFeature_RNA", "nCount_RNA", "percent_mito", "percent_rbp","percent_plt"), ncol = 5) +
         theme(plot.title = element_text(size=15))

```

##### Normalise

```{r normalise,echo=T,include=T, message = FALSE}

srat_data.norm <- NormalizeData(srat_data,
                                normalization.method = "LogNormalize",
                                scale.factor = 10000)


srat_data.norm <- FindVariableFeatures(srat_data.norm,
                                       selection.method = "vst",
                                       nfeatures = 2000)
srat_data.norm <- ScaleData(srat_data.norm)

# devtools::install_github("krlmlr/ulimit")
# setTimeLimit(cpu = Inf, elapsed = Inf, transient = FALSE)
# srat_data.norm <-  SCTransform(srat_data,verbose=TRUE) #--> cpu-time-limit kills the process !!!

## also: mean.function = ExpMean

top20 <- head(VariableFeatures(srat_data.norm), 20)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(srat_data.norm)

LabelPoints(plot=plot1, points = top20, repel = TRUE, xnudge=0, ynudge=0)

```

##### Perform linear dimensional reduction

```{r pca,echo=T,message=F}

srat_data.norm <- RunPCA(srat_data.norm,
                         features = VariableFeatures(object = srat_data.norm))

ElbowPlot(srat_data.norm)

DimPlot(srat_data.norm, reduction = "pca")

```

##### CLUSTER

```{r cluster,echo=T,include=T }

srat_data.norm <- FindNeighbors(srat_data.norm, dims = 1:30)

srat_data.norm <- FindClusters(srat_data.norm, resolution = 0.4)

##>Run non-linear dimensional reduction  
srat_data.norm <- RunUMAP(srat_data.norm,
                          dims = 1:30,
                          n.components = 3L,
                          verbose=TRUE)

srat_data.norm %>% glimpse

DimPlot(srat_data.norm, reduction = "umap",label=TRUE)

```

##### Annotation of clusters

```{r define_clusters,echo=T,include=T}

srat_data.norm <- SetIdent(srat_data.norm, value = "cell_annotation")

Idents(srat_data.norm)[1:10]


#####> If we want to define the cluster-names on our own :
# srat_data.norm <- RenameIdents(object= srat_data.norm,
#                                "0"= , "1"= , "2"=,..)

DimPlot(object = srat_data.norm,
        reduction="umap",  
        pt.size = 0.5,
        label.size = 3,label = TRUE,label.box = TRUE)+
        theme(legend.position = "none")
```


```{r ditto_DimPlot}

library(dittoSeq)

dittoDimPlot(srat_data.norm,
             var = "ident",
             reduction.use = "umap",
             do.label = T,
             main="Human Muscle Clusters(DeMicheli)")
# var :  	String name of a "gene" or "metadata"(eg."SCT_snn_res.0.4") to use for coloring the plots.

```

```{r gene_markers,echo=T,include=T}
# muscle.markers <- FindAllMarkers(srat_data.norm,
#                                  test.use = "wilcox",
#                                  only.pos = T,
#                                  min.pct = 0.25,
#                                  logfc.threshold = 0.25,
#                                  verbose = TRUE)

##>replaced by :

muscle.markers <- readRDS("../Seurat_Analysis/muscle_markers.rds") #( calculated in "analyse_norm_Data.Rmd")

# muscle.markers %>% pull(cluster) %>% as.factor %>% levels
# muscle.markers %>% group_by(cluster) %>% slice_max(n=2,order_by = avg_log2FC,with_ties=FALSE ) #n
# = show this number of lines per cluster
##> visualise per marker:
##> RidgePlot, CellScatter, and DotPlot as additional methods to view your dataset
cat("\nCurrent cluster ids = ", paste((muscle.markers$cluster %>% as.factor %>% levels),collapse="\n"))


```


###### Identify clusters by TOP-10 genes from Markers

```{r group_by_heatmap, echo=T, eval=T}

top10Up <- muscle.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)

DoHeatmap(srat_data.norm, features = top10Up$gene) + NoLegend()

#pdf("Output/GeneMarkers_Dot_Plot.pdf", width=15, height=15)
DotPlot(srat_data.norm,
        features = as.character(unique(top10Up$gene)),
        assay = "RNA", dot.scale=1) +
        coord_flip() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1,size=9),
              axis.text.y = element_text(angle = 0, hjust = 1,size=5))

#dev.off()


```

####> Identify Fibroblasts-Markers :

```{r}

fibro_Markers <- Seurat::FindMarkers(srat_data.norm, ident.1=c("COL1A1+ Fibroblasts",
                                                          "DCN+ GSN+ MYOC+ Fibroblasts",
                                                          "FBN1+ MFAP5+ CD55+ Fibroblasts"),
                                    logfc.threshold = 0.25,only.pos = T, min.cells.feature = 3)

fibro_Markers$GeneName <- rownames(fibro_Markers)

fibro_Markers$GeneName %>% unique %>% length  ##1151

write_delim(fibro_Markers, file="Fibroblasts_3subtypes_FindMarkers_DeMicheli_hum_logfc-0.25.tsv",
            delim="\t",col_names=T)

```

#### ...or , Each one separetely:

```{r}

fibro_col1a1_markers <- Seurat::FindMarkers(srat_data.norm,
                        ident.1=c("COL1A1+ Fibroblasts"),
                        ident.2 = NULL, logfc.threshold = 0.25,
                        only.pos = T,min.cells.feature = 3)

##!! Attention: geneNames in Rownames -> not saved with "write_delim":
fibro_col1a1_markers$GeneName <- rownames(fibro_col1a1_markers)

fibro_col1a1_markers$GeneName %>% unique %>% length


fibro_myoc_markers <- Seurat::FindMarkers(srat_data.norm,
                       ident.1=c("DCN+ GSN+ MYOC+ Fibroblasts"),
                       ident.2 = NULL, logfc.threshold = 0.25,
                       only.pos = T,min.cells.feature = 3)
fibro_myoc_markers$GeneName <- rownames(fibro_myoc_markers)

fibro_myoc_markers$GeneName %>% unique %>% length

fibro_fbn1_markers <- Seurat::FindMarkers(srat_data.norm,
                      ident.1=c("FBN1+ MFAP5+ CD55+ Fibroblasts"),
                      ident.2 = NULL, logfc.threshold = 0.25,only.pos = T)

fibro_fbn1_markers$GeneName <- rownames(fibro_fbn1_markers)

fibro_fbn1_markers$GeneName %>% unique %>% length


####> Save in file :
dir.create("For_Secretome/",showWarnings = F)

write_delim(fibro_col1a1_markers,"Secretome/markers_of_COL1A1_fibroblasts_hs.tsv",col_names=T,delim="\t")

write_delim(fibro_myoc_markers,"Secretome/markers_of_MYOC_fibroblasts_hs.tsv",col_names=T,delim="\t")

write_delim(fibro_fbn1_markers,"Secretome/markers_of_FBN1_fibroblasts_hs.tsv",col_names=T,delim="\t")


```

article for DE_genes tools : https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7455704/
tutorial : https://hms-dbmi.github.io/scw/differential-expression.html
